Compositional Equivalences Based on Open pNets

### 2022-10-07 2016-Weak-FH-Bisim/WeakBisim.tex
typos

We develop here a semantics for a model of interacting processes with parameters and holes. 
Our approach is originally inspired from Structured Operational Semantics with conditional premisses as in \cite{GROOTE1993263,VANGLABBEEK2004229}.
But we aim at a more constructive and implementable approach to compute the semantics (intuitively transitions including first order predicates) and to 
check equivalences for these open systems.
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures; these properties will then be preserved
by any correct instantiation of the holes.
As a consequence, our model allows us to reason about composition
operators as well as about realistic distributed systems. The parametric
nature of the model and the properties of compositionality of the
equivalence relations are thus  the main
strengths of  our approach.

@@

\item	The  rule \TrDeux\ deals with pNet nodes: for each possible
	synchronisation vector (of index $k$) applicable to the rule subject, the premisses
	include one {\em open transition} for each sub-pNet involved, one possible
	{\em action} for each hole involved, and the predicate relating these
	with the resulting action of the vector. The sub-pNets involved are split between two 
	sets, $I_2$ for sub-pNets that are pLTSs (with open transitions obtained by rule \textbf{Tr1}), and $I_1$ for the sub-pNets that are not pLTSs (with open transitions obtained by rule \TrDeux), $J$ is the set of 
	holes involved in the transition\footnote{Formally, if $SV_k \!=\! \SV{({\alpha'})_m^{m 
	\in M}}{\alpha'}{e_b}$ is a synchronisation vector  of \pNet\  then $J=M\cap 
	\Holes(\pNet)$, $I_2=M\cap \Leaves(\pNet)$,  $I_1=M\setminus J \setminus 
	I_2$}\footnote{We could replace $I_1$ and $I_2$ by their formal definition in \TrDeux\ but the rule would be more difficult to read.}.   
\end{itemize}                                                                  
%\begin{mathpar}

@@

There is also a clear relation with the seminal works on rule formats for Structured Operational Semantics, e.g. DeSimone format, GSOS, and conditional rules with or without negative premisses \cite{deSimone85,Bloom:1995,GROOTE1992202,VANGLABBEEK2004229}.  The Open pNets model provides a way to define operators similar to these rules formats, but with quite different aim and approach. A formal comparison would be interesting, though not trivial. What we can say easily is that: the pNet format syntactically encompasses  De Simone, GSOS, and conditional premisses rules. Then our compositionality result is more powerful than their classical results, but this is not a surprise, as we rely on a (sufficient) syntactic hypothesis on a particular system, rather than the general rules defining an operator.  Last, we intentionally do not accept negative premisses, that would be more  to put into practice in our implementation. This extension could be studied in future work.

>>>>

We develop here a semantics for a model of interacting processes with parameters and holes. 
Our approach is originally inspired from Structured Operational Semantics with conditional premises as in \cite{GROOTE1993263,VANGLABBEEK2004229}.
But we aim at a more constructive and implementable approach to compute the semantics (intuitively transitions including first order predicates) and to 
check equivalences for these open systems.
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures; these properties will then be preserved
by any correct instantiation of the holes.
As a consequence, our model allows us to reason about composition
operators as well as about realistic distributed systems. The parametric
nature of the model and the properties of compositionality of the
equivalence relations are thus  the main
strengths of  our approach.

@@

\item	The  rule \TrDeux\ deals with pNet nodes: for each possible
	synchronisation vector (of index $k$) applicable to the rule subject, the premises 
	include one {\em open transition} for each sub-pNet involved, one possible
	{\em action} for each hole involved, and the predicate relating these
	with the resulting action of the vector. The sub-pNets involved are split between two 
	sets, $I_2$ for sub-pNets that are pLTSs (with open transitions obtained by rule \textbf{Tr1}), and $I_1$ for the sub-pNets that are not pLTSs (with open transitions obtained by rule \TrDeux), $J$ is the set of 
	holes involved in the transition\footnote{Formally, if $SV_k \!=\! \SV{({\alpha'})_m^{m 
	\in M}}{\alpha'}{e_b}$ is a synchronisation vector  of \pNet\  then $J=M\cap 
	\Holes(\pNet)$, $I_2=M\cap \Leaves(\pNet)$,  $I_1=M\setminus J \setminus 
	I_2$}\footnote{We could replace $I_1$ and $I_2$ by their formal definition in \TrDeux\ but the rule would be more difficult to read.}.   
\end{itemize}                                                                  
%\begin{mathpar}

@@

There is also a clear relation with the seminal works on rule formats for Structured Operational Semantics, e.g. DeSimone format, GSOS, and conditional rules with or without negative premises \cite{deSimone85,Bloom:1995,GROOTE1992202,VANGLABBEEK2004229}.  The Open pNets model provides a way to define operators similar to these rules formats, but with quite different aim and approach. A formal comparison would be interesting, though not trivial. What we can say easily is that: the pNet format syntactically encompasses  De Simone, GSOS, and conditional premises rules. Then our compositionality result is more powerful than their classical results, but this is not a surprise, as we rely on a (sufficient) syntactic hypothesis on a particular system, rather than the general rules defining an operator.  Last, we intentionally do not accept negative premises, that would be more  to put into practice in our implementation. This extension could be studied in future work.
